---
title: "Traffic fatalities"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---


```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(janitor)
library(readxl)
library(lubridate)
```

## Import 

- FARS 2018 data from [USDOT FARS database](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars). Bulk download is [via ftp](ftp://ftp.nhtsa.dot.gov/fars/).
- Cleaned state/county values is from previous notebook "01-glc-import-clean.Rmd".

```{r}
accidents_2018 <- read_csv("data-raw/FARS2018NationalCSV/ACCIDENT.csv") %>% clean_names()
glc_state_county <- read_rds("data-out/01-glc_state_county.rds")
```

### Change to text

Some columns came in as numbers that should be text. These are mostly codes that need joining.

```{r}
accidents_2018 <- accidents_2018 %>% 
  mutate(
    county = county %>% as.character(),
    state = state %>% as.character(),
    city = city %>% as.character(),
    st_case = st_case %>% as.character()
    )
```

### Create a date column

```{r}
accidents_2018 <- accidents_2018 %>% 
  mutate(
    acc_date = ymd(paste(year, month, day, sep = '-'))
    )
```


### Join to get geography names

The accidents uses codes for state and county. We use the GLC data to get those.

```{r}
accidents_2018 <- accidents_2018 %>% 
  left_join(glc_state_county, by=c("state" = "state_code", "county" = "county_code"))
```

#### Testing for misspelled county names

When joining these files, I found I got more observations than the original file. This was because there were multiple spellings for certain counties.

Create a dataframe that is the count of multiple cases. This would happen if there were multiple county_names, which there were, but they hopefully have been fixed in the glc-import-clean notebook.

```{r}
cnty_spell <- accidents_2018 %>% 
  count(st_case) %>% 
  arrange(n %>% desc()) %>% 
  filter(n > 1) %>% 
  pull(st_case) %>% 
  tibble::enframe(name = NULL)

cnty_spell <- cnty_spell %>% 
  rename(st_case = value)
```

Join the accidents to the cnty_spell to find the multiple names. This should show zero rows.

```{r}
accidents_2018 %>% 
  inner_join(cnty_spell) %>% 
  select(state_name, county_name) %>% 
  distinct()
```

## Export data

```{r}
accidents_2018 %>% 
  write_rds("data-out/02_accidents_2018.rds")
```

